HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series

### 2022-11-02 signal_processing_elsevier/re-submission_ijf/Response_to_reviewers/hermes_paper_response_to_reviewers.tex

\documentclass[11pt]{article}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{xcolor}

\title{Cover letter for the revised version of IJF-D-22-00224}

\author{\'Etienne David et al.}
%\date{}                                           % Activate to display a given date or no date

\begin{document}
\maketitle
%\section{}
%\subsection{}

We would like to thank the reviewers and the associate editor for attentively reading our work, the constructive feedback, and their appreciation of the paper. We are also grateful for the fast handling of the manuscript.
We have carefully considered all the comments of the reviewers. All the minor errors and typos listed by the reviewer have been corrected as suggested and you can find detailed responses to the comments below.

\subsection*{Major changes}

The first major change concerns the dataset description in Section 3. In section 3.1, an overview of the image recognition framework is provided with references to the main architectures used. In section 3.2, a more detailed explanation is given on how the images are translated to fashion time series and how they are normalized to remove social media bias. Finally, a clearer description of the external signal is provided in Section 3.3.

The second major change is the update of Section 4.4 concerning the M4 weekly dataset. We have added to the pool of benchmarks the top 3 models attaining the highest accuracy during the M4 competition on the weekly dataset. As these models are ensembling methods, an ensembling approach mixing 4 HERMES variations is proposed. This proposed ensembling achieves the same level of accuracy as the previous top 3 methods and illustrates that HERMES models can be easily introduced in an ensemble framework to provide more robust and accurate final predictions.

In addition to the two main changes, we want to highlight several additional updates. i) Figure 3 is updated to display a clearer big picture of the HERMES framework. ii) In Appendix A.3, a complete description of the proposed ensembling combining HERMES variations is added for completeness. iii) Finally, in Appendix B.2, we describe the entire grid search run on the M4 weekly dataset to set the hyperparameters of the HERMES architecture.

\subsection*{Reviewer 1}

\begin{itemize}
\item {\em ``In page 3, authors claim that they use cutting-edge image recognition techniques to create the "first fashion dataset". However, ... . Also, the target variable to forecast, "share of category" is also not clearly explained or provided the exact equations used to derive these values. ''} \medskip

\textbf{Author response:}  We agree perfectly with the referee and this comment has led to an important update of our paper to better describe the dataset. We have now presented more precisely the creation of the dataset: from the image recognition framework (Section 3.1) to the creation of the fashion time series and the "influencer" external signals (Section 3.2 and 3.3).

\item {\em ``Authors highlight that they avoid using ensemble-based benchmarks on M4 dataset for a fair comparison. However, HERMES also mimics a similar functionality to that of ensemble models (e.g., Boosting methodology found in the ML literature) as it uses a combination of methods.''} \medskip

\textbf{Author response:}  We thank the referee for this sensible suggestion that motivates an improvement of Section 4.4. In addition to the existing comparison between the HERMES model and a pool of benchmarks, a second comparison has been done between the 3 ensembling methods reaching the highest accuracy on the M4 weekly dataset and an ensembling combining 4 HERMES variations. The HERMES hybrid model can easily be included in an ensembling framework and the proposed ensembling method reaches the same accuracy level as the best approaches of the M4 competition.

\item {\em ``Authors must avoid the ambiguity in the paper. For example, in page 12, authors say "existing Python or R libraries are used to estimate the different parameters". Why do you use a "or" in this sentence? Authors must be more specific about what they implement (exact packages etc)''} \medskip

\textbf{Author response:}  We have now corrected parts with ambiguity as suggested.

\item {\em ``Authors use a subset of the fashion time series to demonstrate the use of HERMES on smaller dataset. Again, have not explained the details of the selection of subset of data. Are they randomly selected or handpicked?''} \medskip

\textbf{Author response:} There was indeed an oversight on this part. The two subsets of data are selected randomly and the text has been updated accordingly.

\item {\em ``Authors have used a grid-based methodology to determine the optimal parameters for the neural network-based methods but have not clearly mentioned the initial parameter ranges used in the experiments. In page 27, authors say that "have been trained with a range of values". To reproduce the results, it is important you provide these details (since you don't provide the source code, authors should at least make an effort to give these information)''} \medskip

\textbf{Author response:}  To complete Appendix B.1 where we present a grid search run on the loss function, we have added in Appendix B.2 a complete example of grid search run for all the hyperparameters of the HERMES architecture on the M4 weekly dataset.

\item {\em ``Authors must use consistent benchmarks across both datasets (ES-RNN benchmark missed in the Fashion dataset). Also haven't provided a statistical test to evaluate the significance of the differences of the results (the results are statistically significant or not)''} \medskip

\textbf{Author response:} We agree with the reviewer, the absence of the ES-RNN benchmark on the Fashion dataset is annoying. However, this choice has been motivated for two reasons: Firstly, the ES-RNN code base is developed in a non-usual code framework (Dynet) that could make the migration to Python difficult. A Python package exists but it is especially designed to work on the M4 dataset and we decided not to update the code to the Fashion dataset use case. Secondly, to compensate this absence: i) we partially replicate it with a method called \textit{hermes-ets} on the fashion dataset and ii) we run numerous experiments on the M4 weekly dataset to provide a complete comparison between the true ES-RNN and the HERMES approach.

\item {\em ``In page 15, it is also not clear about the exact classification algorithms used in the study. Authors don't not refer nor provide justification for the use of proposed the threshold (5\%) to label the time series into decreasing/flat/increasing trends. Of course, it helps to create a balance dataset, but that is not a valid justification to use this threshold. Need to import more evidence/justification from the fashion domain''} \medskip

\textbf{Author response:} We totally understand the expectations of the referee with this remark. However, the chosen threshold was selected to match the fashion industry practices. We decided to provide additional numerical simulations, for instance in the M4 setting, to illustrate the performance of the model instead of adding results with other thresholds which would not be used in practice.

\item {\em ``The performance of the RNN on the M4 weekly time series dataset seems to be poor. Given the proposed HERMES also use this algorithm (Figure 6) as a part of the framework, it is a bit surprising to see how it improves the accuracy after the inclusion of per series TBATS model (Hermes-tbats). I can understand why RNN model is underperforming (because the time series are heterogenous with different starting and end date, and if you globally train the model, without additional/external information, it will not pick up the correct signals).''} \medskip

\textbf{Author response:} We agreed with the reviewer that the M4 weekly dataset is a particularly challenging dataset for RNN approaches. For the RNN benchmark, the neural network has to rebuild the entire prediction while in the HERMES framework, the RNN part task is only to correct an already computed prediction. Thus, the final accuracy of the 4 HERMES variations is mainly supported by their collections of  per-time-series predictors. To illustrate this point, we have added Table A.7 in Appendix A which shows the accuracy of the predictors alone and included in the HERMES framework on the M4 weekly dataset. For each version, the RNN corrector always increases even slightly the final overall accuracy. Moreover, several interesting corrections of the RNN part can be noted as in Figure A.9.

\end{itemize}
\subsection*{Reviewer 2}
\begin{itemize}

\item {\em ``On p.7, in the formula above eq.1 It appears that z is centered around Ypred[T+k], with k=i-h[i/j] . I assume that i/j is an integer division. This is a strange formula, where h seems to be assumed to be equal seasonality. Also, why would you center the input by the output prediction value?''} \medskip

\textbf{Author response:} We thank the referee for this remark and we have now improved this part to give more motivation about our preprocessing. We center the input by the output prediction value for two reasons: i) to include the per-time-series prediction in the RNN input ii) to remove from the RNN input the fundamental patterns already learned by the per-time-series predictor.

\item {\em ``On p.13, fig.6 What is the meaning of e.g. (None,52,50) or (None,50)?''} \medskip

\textbf{Author response:} There was indeed a mistake in the figure 6, (None,52,50) and (None,50) are displayed in the graph without additional information. In order to shed more light on this point, we have decided to merge the RNN part architure and the overall HERMES architecture in a same figure (Figure 3). Information about the size of the LSTM and Dense layer are no longer represented but can be found in Section 4.1.

\item {\em ``On p.14, line 203 Why limit yourself to only one version of ETS (additive, seasonal, no trend I assume)''} \medskip

\textbf{Author response:} We totally agree with the referee. We have now introduced two different versions of the ETS model on the M4 weekly dataset. The first one is an ETS model with an additive trend, seasonality and noise and the second one is the multiplicative ETS with a multiplicative trend, seasonality and noise.

\item {\em ``On p. 18, lines 255-260 I would assume that reducing the dataset size 10 times, would necessitate an update the model/training hyperparameters. No?''} \medskip

\textbf{Author response:} We thank the reviewer for this sensible suggestion. In fact, for each new use case or experiment, a grid search can be run to improve the final accuracy of the HERMES architecture. However, we have noticed that the hyperparameters, except for the choice of the learning rate, do not impact significantly the final accuracy of the model. It is why we have decided not to change the hyperparameters for the experiments on the subsamples of the fashion dataset and also on the M4 weekly dataset, where the same hyperparameters are used for all the HERMES variations.

\item {\em ``On p.2, line 302; Table 5 You mention that a model called "Uber" is added. But I do not see it in table 5. Instead there is there a strange phrase "S.Smyl Hyndman et al. (2020)"''} \medskip

\textbf{Author response:} We have now corrected this
mistake.

\item {\em ``General remark: You fit an NN model on residuals from a tbats model. But tbats has some subversions, e.g. may or may not use Box-Cox transformation. I would suggest that passing to NN the fitted tbats model version could be useful.''} \medskip

\textbf{Author response:} In the hybrid model proposed by S.Smyl, the per-time-series ETS models are not fitted and the RNN part has consequently two tasks, set the parameters of the ETS models depending on the time series and correct their predictions. In the HERMES framework, the per-time-series predictor is first fitted using a Python package and then passed to the RNN corrector. This allows the proposed hybrid model to be built with any kind of per-time-series predictors, even complex one, depending on the use case.
\end{itemize}

\subsection*{Reviewer 3}
\begin{itemize}
\item {\em ``The paper gives the out of sample test results for all horizons, h=1, 2, …52 weeks. It is also interesting to see the performance of the methods separately for short (one week), medium (one month) and long term (one year) horizons.''} \medskip

\textbf{Author response:} With the applications on the Fashion dataset and the M4 weekly dataset, the HERMES approach has been tested with two different horizons: h=13 and h=52. Nevertheless, looking at the impact of the Influencer signal on the main signal (as displayed in Figure 1), it would be interesting to focus on a short term forecast. This point is not developed in this paper but is currently fueling our research work.

\item {\em ``Explain more clearly how you have constructed the values of weak signals.''} \medskip

\textbf{Author response:}  We agree perfectly with the referee and this comment has led to a major update of our paper. We have now presented more precisely the creation of the fashion time series and the "influencer" external signal (Section 3.2 and 3.3).

\item {\em `` Page 12: Explain why you have selected the window size as 104, also the same for M4 data. Are the results sensitive to window size?''} \medskip

\textbf{Author response:} So as to highlight the impact of the different hyperparameters of the HERMES architecture on the final accuracy, we have run a complete grid search on the M4 weekly data and displayed the result in Appendix B. The window size is a determinant parameter and we have noticed that giving at least two seasonal periods to the RNN corrector usually brings better results on average.

\item {\em ``Page 16, What is the Accuracy metric in table 2? Is it percentage of correct classification?''} \medskip

\textbf{Author response:} We have added a sentence at the end of Section 4.2 to describe the Accuracy metric.

\item {\em ``Page 18: Did you select your 1000 and 100 time series randomly?''} \medskip

\textbf{Author response:} There was indeed a oversight on this part. The two subsets of data are selected randomly and the text has been updated

\item {\em ``Write the definition of OWA.''} \medskip

\textbf{Author response:} We agree with the referee and we have added a section describing the OWA in Appendix A.

\item {\em ``The paper has some spelling mistakes and inconsistencies, e.g. in somes cases (Page 6) and SMAE and MASE (used in page 26). Correct the spelling and use the notations consistently.''} \medskip

\textbf{Author response:} We thank the reviewer and the text has been corrected accordingly.
\end{itemize}
\end{document}

### 2022-11-02 signal_processing_elsevier/re-submission_ijf/hermes_paper_LaTeX_source_files/hermes_paper_elsevier.tex
re-submission hermes

\usepackage{lineno,hyperref}
%\usepackage[utf8]{inputenc}
%\usepackage[english]{babel}
%\usepackage{authblk}
\usepackage{amssymb,amsmath,amsthm}
%\usepackage[pdftex]{graphicx}
%\usepackage{footnote}
%\usepackage{url}
%\usepackage{times,dsfont}
%\usepackage{fancyhdr}
%\usepackage{geometry}
%\pagestyle{fancy}
\modulolinenumbers[5]

@@

%% Group authors per affiliation:
\author[mymainaddress,mysecondaryaddress]{Etienne DAVID}

@@

\begin{abstract}
Developing models and algorithms to draw causal inference for time series is a long standing statistical problem. It is crucial for many applications, in particular for fashion or retail industries, to make optimal inventory decisions and avoid massive wastes. By tracking thousands of fashion trends on social media with state-of-the-art computer vision approaches, we propose a new model for fashion time series forecasting. Our contribution is  twofold. We first provide publicly\footnote[1]{\url{http://files.heuritech.com/raw_files/f1_fashion_dataset.tar.xz}} the first fashion dataset gathering \numberts\ weekly fashion time series. As influence dynamics are the key of emerging trend detection, we associate with each time series an external weak signal representing behaviors of influencers. Secondly, to leverage such a complex and rich dataset, we propose a new hybrid forecasting model. Our approach combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. This hybrid model provides state-of-the-art results on the proposed fashion dataset, on the weekly time series of the M4 competition \cite{makridakis2018m4}, and illustrates the benefit of the contribution of external weak signals.
\end{abstract}

@@

Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \cite{sarkka2013bayesian, douc2014nonlinear, zucchini2017hidden} and the numerous references therein.
 %Due to the diversity of applications and use cases, a multitude of models have been proposed. 
Parametric generative models provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations. %The parameters of these models are usually estimated using a sequence of observations of the target time series. 
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \cite{RePEc:inm:oropre:v:9:y:1961:i:5:p:673-685}, the Trigonometric Box-Cox transform, ARMA errors, Trend, and Seasonal components model (TBATS) \cite{doi:10.1198/jasa.2011.tm09771}, or the ARIMA with the Box-Jenkins approach \cite{box2015time} are for instance very popular parametric generative models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain, the observations arising therefrom have a complex statistical structure. %Introducing latent hidden states, they assume that the predictive distribution is not unique and constant in time but multiple and lead by the values of the hidden states changing during time. 
In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \cite{touron2017modeling}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \cite{touron2019consistency} and Expectation Maximization based numerical experiments were proposed.
Although these works provide promising results, HMM are computationally expensive to train and are not yet well studied for seasonal  sequences with thousands of components.

In many fields, single or few time series have become thousands of sequences with various statistical properties. In this new context, classical time series specific statistical models show limitations when dealing with numerous heterogeneous data. Recurrent neural networks and recent sequence to sequence deep learning architectures offer very appealing numerical alternatives thanks to their capability of leveraging any kind of heterogeneous multivariate data, see for instance \cite{ hochreiter1997long,vaswani2017attention, 8614252, li2019enhancing, lim2019temporal,salinas2020deepar}. The DeepAR model proposed in \cite{salinas2020deepar} provides a global model from many time series based on a multi-layer recurrent neural network with LSTM cells. More recently, applications using the Transformer model have been proposed  \cite{li2019enhancing}. The Temporal Fusion Transformers (TFT) approach is a direct alternative to the DeepAR model \cite{lim2019temporal}.  Unfortunately, all these solutions suffer from two main weaknesses. Firstly, many of them are black-boxes  as the final forecast usually does not come with a statistical guarantee  although a few recent works focused on measuring uncertainty in recurrent neural networks, see  \cite{martin2020monte}. Secondly, without a fine preprocessing and well chosen hyper-parameters, these methods may lead to poor results and be outperformed by traditional statistical models, see \cite{makridakis2018m4}.

In this paper, we consider a  new time series forecasting application referred to as {\em fashion trends prediction}. Based on a cutting-edge image recognition technology, we built the first fashion dataset containing \numberts\ weekly sequences  of fashion trends on social media  from 01-01-2015 to 01-01-2019. This dataset has very appealing properties:  all time series have the same length, no missing value and there is no sparse time series even for niche trends. The originality of our dataset comes from the fact that additional external weak signals can also be introduced. With our fashion expertise, we detected several groups of highly influential fashion users. Analyzing their specific behaviours on social media, we associate with each time series an external weak signal representing the same fashion trends on a sub-category of users. They are called weak signals because they are often alerts or events that are too sparse, or too incomplete to allow on their own an accurate estimation of their impact on the prediction of the target signal. With this totally new application, we aim at designing a model able to deal with such a large dataset, leveraging  complex external weak signals and finally providing the most accurate forecasts.

Recurrent neural networks are appealing to tackle our forecasting problem due to their capability of leveraging external data.  Recently, hybrid models combining deep neural network (DNN) architectures with widespread statistical models to deal with seasonality and trends have been proposed, see for instance  \cite{zhang2003time,jianwei2019novel,bandara2020lstm}. The approach providing the most striking results was proposed in  \cite{smyl2020hybrid} in the context of the M4 forecasting competition \cite{makridakis2020m4}.  Given a large dataset, a per-time-series multiplicative exponential smoothing model was introduced to estimate simple but fundamental components for each time series and compute a first prediction. Then a global recurrent neural network was trained on the entire dataset to correct errors of the previous exponential smoothing models.

@@

@@

\subsection{Per-time-series predictors}
The time-series-specific predictors compute, for each sequence, a first $\lag$-ahead prediction based on the past. For all $1\leqslant n \leqslant N$, we note $\stat^n(.;\statparam^n)$ the $n$-th parametric model of the $n$-th sequence where $\statparam^n$ are  unknown parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$ and the estimated  parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{pred,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{pred,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i\,.
$$
During the M4 competition, the hybrid model of \cite{smyl2020hybrid} was based on a multiplicative exponential smoothing model as the time-series-specific predictor. However, on sporadic time series, this choice leads to poor results and instability. In this paper, a more general framework able to deal with any kind of per-time-series models is provided. In Section~\ref{sec:exp}, two versions of our framework are proposed. The first one is based on an  exponential smoothing as a reference similar to the baseline \cite{smyl2020hybrid} and the second one uses a TBATS model \cite{doi:10.1198/jasa.2011.tm09771} which provides better results as this parametric model includes  Fourier representations with time varying coefficients, and ARMA error correction.

In the specific application of fashion trends prediction, two main forms of standard weaknesses were detected and defined for univariate parametric forecasts. The first one is a weakness of definition, when the model is not well defined for its forecasting task. For instance, an additive model provides good predictions on time series with additive seasonality but fails at handling multiplicative seasonalities. The second form, harder to detect and correct, is the weakness of information. For non stationary time series, huge changes of behaviours are not always predictable using the past of the sequence. In somes cases, these changes depend on external variables not considered by univariate parametric models. The difficulty is that the exact influence of external variables on the main signal is mostly unknown. This motivates the introduction of a global RNN trained on all time series and able to consider and leverage external signals. First forecast example of our new hybrid model is given in Figure~\ref{fig:introexamples}.

\begin{figure*}
\centering
%\begin{subfigure}{1.\textwidth}
 % \centering
  %\includegraphics[width=\linewidth]{figure/us_female_shoes}
  %\caption{Time series representing a shoes fashion trend for female in The United States.}
 % \label{fig:introexamples:sub1}
%\end{subfigure}
%\begin{subfigure}{1.\textwidth}
  %\centering
  \includegraphics[width=\linewidth]{figure/br_female_texture_verticalstripe}
 % \caption{Time series representing the vertical stipes texture fashion trend for female in Brazil.}
  %\label{fig:introexamples:sub2}
%\end{subfigure}
\caption{Hermes forecast examples. In green the prediction of the TBATS per-time-series predictors. In red the final forecast of our HERMES hybrid model.  Time series representing the vertical stipes texture fashion trend for females in Brazil.}%(Top) Time series representing a shoes fashion trend for females in The United States. (Bottom) Time series representing the vertical stipes texture fashion trend for females in Brazil.}
\label{fig:introexamples}
\end{figure*}

@@

Let $\window \in \mathbb{N}$ be the window size, usually this window is proportional to the forecast horizon $\window \propto \lag$. The RNN input is defined as the following  normalized, deseasonalized and rescaled sequence $\rnnwindow^n_T = \{\rnninput^{n}_{T-\window+i|T}\}_{1\leqslant i \leqslant w}$, where, for all $1\leqslant n \leqslant N$, $1\leqslant i \leqslant w$ and $k = i - h\lfloor i/h \rfloor$, 
$$
\rnninput^{n,T}_{T-w+i|T} = \frac{\ts^n_{T-w+i} -\tspred^{pred,n}_{T+k|T}}{\meants^n_T}\,,\quad \meants^n_T = \frac{1}{w}\sum_{i = 1}^{w}\ts^n_{T-w+i}\,.
$$
Let $\rnn(.;\rnnparam)$ be the recurrent neural network model where $\rnnparam$ are  unknown parameters. Given the RNN input sequences $\{\rnnwindow^n_T\}_{1\leqslant n \leqslant N}$ and the global RNN estimated parameters $\rnnparam$, the error-corrector predictions $\{\tspred^{corr,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{corr,n}_{T+i|T} = \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\ \,.
$$
Our hybrid model forecast is, for all $1\leqslant n \leqslant N$ and all $i \in \{1,\ldots,\lag\}$,
\begin{align}
\label{eq:nows:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T} \\
&= \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}

>>>>

\usepackage{lineno,hyperref}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{xcolor}
\usepackage{pdfpages}
\modulolinenumbers[5]

@@

%% Group authors per affiliation:
\author[mymainaddress,mysecondaryaddress]{\'Etienne DAVID}

@@

\begin{abstract}
Developing models and algorithms to draw causal inference for time series is a long standing statistical problem. It is crucial for many applications, in particular for fashion or retail industries, to make optimal inventory decisions and avoid massive wastes. By tracking thousands of fashion trends on social media with state-of-the-art computer vision approaches, we propose a new model for fashion time series forecasting. Our contribution is  twofold. We first provide publicly\footnote[1]{\url{http://files.heuritech.com/raw_files/f1_fashion_dataset.tar.xz}} an appealing fashion dataset gathering \numberts\ weekly fashion time series. As influence dynamics are the key of emerging trend detection, we associate with each time series an external weak signal representing behaviors of influencers. Secondly, to leverage such a complex and rich dataset, we propose a new hybrid forecasting model. Our approach combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. This hybrid model provides state-of-the-art results on the proposed fashion dataset, on the weekly time series of the M4 competition \cite{makridakis2018}, and illustrates the benefit of the contribution of external weak signals.
\end{abstract}

@@

Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \cite{sarkka2013, douc2014, zucchini2017} and the numerous references therein.
Parametric generative models provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations.
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \cite{Brown1961}, the Trigonometric Box-Cox transform, ARMA errors, Trend, and Seasonal components model (TBATS) \cite{alysha2011}, or the ARIMA with the Box-Jenkins approach \cite{box2015} are for instance very popular parametric generative models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain, the observations arising therefrom have a complex statistical structure.
In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \cite{touron2017}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \cite{touron2019} and Expectation Maximization based numerical experiments were proposed.
Although these works provide promising results, HMM are computationally expensive to train and are not yet well studied for seasonal  sequences with thousands of components.

In many fields, single or few time series have become thousands of sequences with various statistical properties. In this new context, classical time series specific statistical models show limitations when dealing with numerous heterogeneous data. Recurrent neural networks and recent sequence to sequence deep learning architectures offer very appealing numerical alternatives thanks to their capability of leveraging any kind of heterogeneous multivariate data, see for instance \cite{ hochreiter1997,vaswani2017, siami2018, li2019, lim2019,salinas2020}. The DeepAR model proposed in \cite{salinas2020} provides a global model from many time series based on a multi-layer recurrent neural network with LSTM cells. More recently, applications using the Transformer model have been proposed  \cite{li2019}. The Temporal Fusion Transformers (TFT) approach is a direct alternative to the DeepAR model \cite{lim2019}.  Unfortunately, all these solutions suffer from two main weaknesses. Firstly, many of them are black-boxes  as the final forecast usually does not come with a statistical guarantee  although a few recent works focused on measuring uncertainty in recurrent neural networks, see  \cite{martin2020}. Secondly, without a fine preprocessing and well chosen hyperparameters, these methods may lead to poor results and be outperformed by traditional statistical models, see \cite{makridakis2018}.

In this paper, we consider an emerging time series forecasting application referred to as {\em fashion trends prediction}. In fashion and retails industries, accurately anticipating consumers' needs is vital and wrong decisions can lead to massive wastes. Analysing social media pictures to detect what clothes people are wearing, it is possible to translate the visibility of fashion items over time into time series. Consequently, models and algorithms can be trained to accurately anticipate and predict consumers' behaviours. In \cite{ma2020}, a  dataset is provided using social media pictures and an image recognition framework to detect several clothes: 2000 fashion time series are proposed with a weekly seasonality. However, only 3 years of historical data is available (144 data points) that may not be sufficient for some statistical approaches. In \cite{ma2020}, another dataset is presented gathering 8000 fashion sequences with an historical available data  increased to 5 years. Nevertheless, only 120 values are available for each fashion time series and the overall volume remains low for a large part of the sequences resulting in a lot of noise and no clear patterns. In this paper, we propose a new fashion dataset overcoming the weaknesses of the two previous ones. Based on cutting-edge image recognition algorithms \cite{ren2015,chollet2017}, we built a large fashion dataset containing \numberts\ weekly sequences of fashion trends on social media with 5 years of historical data from 01-01-2015 to 30-12-2019. This dataset has very appealing properties:  all time series have the same length (261 data points), no missing value and no sparse time series even for niche trends. Concerning fashion dynamics, some of them appear to be really volatile with nonlinear changes of dynamics resulting from the emergence of new tendencies. In this context, understanding early signals of the apparition of a trend is one of the key to accurately forecast the future of the fashion. Consequently, the originality of our dataset comes from the fact that additional external weak signals are introduced. With our fashion expertise, we detected several groups of highly influential fashion users. Analyzing their specific behaviours on social media, we associate with each time series an external weak signal representing the same fashion trends on a sub-category of users. They are called weak signals because they are often alerts or events that are too sparse, or too incomplete to allow on their own an accurate estimation of their impact on the prediction of the target signal. Figure \ref{fig:oneemergingtrend} shows an example of fashion time series with its linked influencers weak signal. Exploring this new way of representing fashion, we aim at designing a model able to deal with such a large dataset, leveraging  complex external weak signals and finally providing the most accurate forecasts.

Recurrent neural networks are appealing to tackle our forecasting problem due to their capability of leveraging external data.  Recently, hybrid models combining deep neural network (DNN) architectures with widespread statistical models to deal with seasonality and trends have been proposed, see for instance  \cite{zhang2003,jianwei2019,bandara2020}. The approach providing the most striking results was proposed in  \cite{smyl2020} in the context of the M4 forecasting competition \cite{makridakis2020}.  Given a large dataset, a per-time-series multiplicative exponential smoothing model was introduced to estimate simple but fundamental components for each time series and compute a first prediction. Then a global recurrent neural network was trained on the entire dataset to correct errors of the previous exponential smoothing models.

@@

\begin{figure*}
  \centering
    \includegraphics[width=1.\linewidth]{br_female_shoes_262}
  \caption{A shoes trend of the fashion dataset. In black the main signal and in orange its associated \textit{fashion-forward} weak signal. The sudden explosion of the influencers signal at the end of 2018 announces the future burst of the trend in the mass market.}
\label{fig:oneemergingtrend}
\end{figure*}

@@

\subsection{Per-time-series predictors}
The time-series-specific predictors compute, for each sequence, a first $\lag$-ahead prediction based on the past. For all $1\leqslant n \leqslant N$, we note $\stat^n(.;\statparam^n)$ the $n$-th parametric model of the $n$-th sequence where $\statparam^n$ are  unknown parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$ and the estimated  parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{pred,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
\begin{equation}
    \label{eq:predictors}
    \tspred^{pred,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i\,.
\end{equation}
During the M4 competition, the hybrid model of \cite{smyl2020} was based on a multiplicative exponential smoothing model as the time-series-specific predictor. However, on sporadic time series, this choice leads to poor results and instability. In this paper, a more general framework able to deal with any kind of per-time-series models is provided. In Section~\ref{sec:exp}, two versions of our framework are proposed. The first one is based on an  exponential smoothing as a reference similar to the baseline \cite{smyl2020} and the second one uses a TBATS model \cite{alysha2011} which provides better results as this parametric model includes  Fourier representations with time varying coefficients, and ARMA error correction.

In the specific application of fashion trends prediction, two main forms of standard weaknesses were detected and defined for univariate parametric forecasts. The first one is a weakness of definition, when the model is not well defined for its forecasting task. For instance, an additive model provides good predictions on time series with additive seasonality but fails at handling multiplicative seasonalities. The second form, harder to detect and correct, is the weakness of information. For non stationary time series, huge changes of behaviours are not always predictable using the past of the sequence. In some cases, these changes depend on external variables not considered by univariate parametric models. The difficulty is that the exact influence of external variables on the main signal is mostly unknown. This motivates the introduction of a global RNN trained on all time series and able to consider and leverage external signals. First forecast example of our new hybrid model is given in Figure~\ref{fig:introexamples}.

\begin{figure*}
\centering
  \includegraphics[width=\linewidth]{br_female_texture_verticalstripe}
\caption{Hermes forecast example on a time series representing the vertical stipes texture fashion trend for females in Brazil. In green the prediction of the TBATS per-time-series predictors. In red the final forecast of our HERMES hybrid model.}
\label{fig:introexamples}
\end{figure*}

@@

Let $\window \in \mathbb{N}$ be the window size, usually this window is proportional to the forecast horizon $\window \propto \lag$. The RNN input is defined as the following  normalized, deseasonalized and rescaled sequence $\rnnwindow^n_T = \{\rnninput^{n}_{T-\window+i|T}\}_{1\leqslant i \leqslant w}$: for all $1\leqslant n \leqslant N$ and $1\leqslant i \leqslant w$,
$$
\rnninput^{n,T}_{T-w+i|T} := \frac{\ts^n_{T-w+i} -\tspred^{pred,n}_{T+k|T}}{\meants^n_T}\,,\quad \meants^n_T = \frac{1}{w}\sum_{i = 1}^{w}\ts^n_{T-w+i}\,.
$$
where $k = i - h\lfloor i/h \rfloor$ with $\lfloor. \rfloor$ the floor function. With the numerator part $\ts^n_{T-w+i} -\tspred^{pred,n}_{T+k|T}$, the per-time-series prediction is included in the RNN input and all the fundamental patterns already learned by this first predictor are removed from the time series. Then the denominator $\meants^n_T$ is use to rescaled all input at the same level as the time series can have different scales. Another option could have been to divide directly  $\ts^n_{T-w+i}$ by $\tspred^{pred,n}_{T+k|T}$ but with time series hitting 0, this option is not valid. Let $\rnn(.;\rnnparam)$ be the recurrent neural network model where $\rnnparam$ are  unknown parameters. Given the RNN input sequences $\{\rnnwindow^n_T\}_{1\leqslant n \leqslant N}$ and the global RNN estimated parameters $\rnnparam$, the error-corrector predictions $\{\tspred^{corr,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{corr,n}_{T+i|T} = \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\ \,.
$$
Our hybrid model forecast is, for all $1\leqslant n \leqslant N$ and all $i \in \{1,\ldots,\lag\}$,
\begin{align}
\label{eq:nows:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T} \\
&= \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}
